# logolink <a href = "https://danielvartan.github.io/brandr/"><img src = "man/figures/logo.svg" align="right" width="120" /></a>

<!-- quarto render -->

```{r}
#| label: setup
#| include: false

library(logolink)

library(badger)
library(brandr)
library(dplyr)
library(ggplot2)
library(here)
library(readr)
library(rutils) # https://github.com/danielvartan/rutils

source(here("R", ".setup.R"))
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badge_repostatus("active"),
  badge_cran_release(),
  badge_github_actions(action = "R-CMD-check.yaml"),
  badge_codecov() |> shush(),
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  badge_custom(
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```
<!-- badges: end -->

## Overview

`logolink` is an [R](https://www.r-project.org/) package that provides simple tools to set up and run [NetLogo](https://www.netlogo.org/) simulations directly from R. It is designed for NetLogo 7 and does not support older versions.

The package offers a modern and streamlined approach to running NetLogo models from R. It follows the [tidyverse
principles](https://tidyverse.tidyverse.org/articles/manifesto.html) and integrates smoothly with the [tidyverse ecosystem](https://www.tidyverse.org/).

> If you find this project useful, please consider giving it a star! &nbsp; [![GitHub repo stars](https://img.shields.io/github/stars/danielvartan/logolink)](https://github.com/danielvartan/logolink/)

## Another R Package for NetLogo?

Yes! Other R packages also connect R and NetLogo, but `logolink` is actively maintained, consistent with tidyverse conventions, and straightforward to use. It is designed to simplify experiment setup and execution while remaining flexible.

For context, [`RNetLogo`](https://cran.r-project.org/web/packages/RNetLogo/index.html) supports older NetLogo versions but is no longer actively maintained. [`nlrx`](https://cran.r-project.org/web/packages/nlrx/index.html) provides a powerful framework for managing experiments and results, but currently [supports](https://docs.ropensci.org/nlrx/reference/supported_netlogo_versions.html) only up to NetLogo 6.3.0 and has [many unresolved issues](https://github.com/ropensci/nlrx/issues). `logolink` complements these packages by focusing on simplicity, full compatibility with NetLogo 7, and seamless integration into modern R workflows.

## Installation

You can install the development version of `logolink` from [GitHub](https://github.com/) with the [`remotes`](https://github.com/r-lib/remotes) package:

``` r
# install.packages("remotes")
remotes::install_github("danielvartan/logolink")
```

## Usage

`logolink` usage is very straightforward. The main functions are:

- [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html): Creates a NetLogo BehaviorSpace experiment XML file
- [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html): Runs a NetLogo BehaviorSpace experiment

Along with this package, you will also need NetLogo 7 or higher installed on your computer. You can download it from the [NetLogo website](https://ccl.northwestern.edu/netlogo/download.shtml).

### Setting the NetLogo Path

`logolink` requires the path to the NetLogo executable when running simulations with the [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html) function. This path is OS-independent, but easy to locate. On Windows, for example, it is typically something like `C:\Program Files\NetLogo 7.0.0\NetLogo.exe`.

Example (Linux):

```{r}
netlogo_path <- file.path("", "opt", "netlogo-7-0-0", "bin", "NetLogo")

netlogo_path
```

### Creating an Experiment

To start running your model from R you first need to setup an experiment. You can do this by setting a [BehaviorSpace](https://docs.netlogo.org/behaviorspace.html) experiment with the [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html) function. This function will create a [XML](https://en.wikipedia.org/wiki/XML) file that contains all the information about your experiment, including the parameters to vary, the metrics to collect, and the number of runs to perform.

Alternatively, you can set up your experiment directly in NetLogo and save it as part of your model. In this case, you can skip the [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html) step and just provide the name of the experiment when running the model with [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html).

Example:

```{r}
library(logolink)

setup_file <- create_experiment(
  name = "Wolf Sheep Simple Model Analysis",
  repetitions = 10,
  sequential_run_order = TRUE,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 1000,
  metrics = c(
    'count wolves',
    'count sheep'
  ),
  run_metrics_condition = NULL,
  constants = list(
    "number-of-sheep" = 500,
    "number-of-wolves" = list(
      first = 5,
      step = 1,
      last = 15
    ),
    "movement-cost" = 0.5,
    "grass-regrowth-rate" = 0.3,
    "energy-gain-from-grass" = 2,
    "energy-gain-from-sheep" = 5
  )
)
```

```{r}
setup_file |> inspect_experiment_file()
```

### Running an Experiment

With the experiment file created, you can now run your model using the [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html) function. This function will execute the NetLogo model with the specified parameters and return the results as a tidy data frame.

```{r}
model_path <- file.path(
  "", "opt", "netlogo-7-0-0", "models", "IABM Textbook", "chapter 4",
  "Wolf Sheep Simple 5.nlogox"
)
```

```{r}
results <- run_experiment(
  netlogo_path = netlogo_path,
  model_path = model_path,
  setup_file = setup_file
)
```

```{r}
library(dplyr)

results |> glimpse()
```

### Analyzing the Data (Bonus Section)

Below is a simple example of how to visualize the results using [`ggplot2`](https://ggplot2.tidyverse.org/).

```{r}
#| output: false

library(dplyr)

data <-
  results |>
  group_by(step, number_of_wolves) |>
  summarise(
    across(everything(), ~ mean(.x, na.rm = TRUE))
  ) |>
  arrange(number_of_wolves, step)
```

```{r}
#| label: readme-wolf-sheep-model-plot

library(ggplot2)

data |>
  mutate(number_of_wolves = as.factor(number_of_wolves)) |>
  ggplot(
    aes(
      x = step,
      y = count_sheep,
      group = number_of_wolves,
      color = number_of_wolves
    )
  ) +
    labs(
      x = "Time step",
      y = "Average number of sheep",
      color = "Wolves"
    ) +
  geom_line()
```

Please refer to the [BehaviorSpace Guide](https://docs.netlogo.org/behaviorspace.html) for complete guidance on how to set and run experiments in NetLogo. To gain a better understand of the basic mechanics behind R and NetLogo communication, see the [Running from the Command Line](https://docs.netlogo.org/behaviorspace.html#running-from-the-command-line) section.

Click [here](https://danielvartan.github.io/logolink/reference/) to see `logolink` full list of functions.

## Citation

If you use this package in your research, please cite it to acknowledge the effort put into its development and maintenance. Your citation helps support its continued improvement.

```{r}
citation("logolink")
```

## License

```{r}
#| echo: false
#| output: asis

cat(
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  )
)
```

```text
Copyright (C) 2025 Daniel Vartanian

logolink is free software: you can redistribute it and/or modify it under the
terms of the GNU General Public License as published by the Free Software
Foundation, either version 3 of the License, or (at your option) any later
version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along with
this program. If not, see <https://www.gnu.org/licenses/>.
```

## Contributing

```{r}
#| echo: false
#| output: asis

cat(
  badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```

Contributions are welcome! Whether you want to report bugs, suggest features, or improve the code or documentation, your input is highly valued. Please check the [issues tab](https://github.com/danielvartan/logolink/issues) for existing issues or to open a new one.

[![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

You can also support the development of `logolink` by becoming a sponsor. Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please mention `logolink` in your donation message.

## Acknowledgments

`logolink` brand identity is based on [NetLogo](https://www.netlogo.org/) brand identity.
