# logolink <a href = "https://danielvartan.github.io/brandr/"><img src = "man/figures/logo.svg" align="right" width="120" /></a>

<!-- quarto render -->

<!-- Install the package before rendering this file: `devtools::install()` -->

```{r}
#| label: setup
#| include: false

library(logolink)

library(badger)
library(brandr)
library(dplyr)
library(ggplot2)
library(here)
library(readr)
library(rutils) # https://github.com/danielvartan/rutils

badge_codecov() |> invisible() # To avoid unnecessary messages.

source(here("R", ".setup.R"))
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis
#| message: false
#| warning: false

cat(
  badge_repostatus("active"),
  badge_cran_release(),
  badge_custom(
    x = "doi",
    y = "10.32614/CRAN.package.logolink",
    color = "1284C5",
    url = "https://doi.org/10.32614/CRAN.package.logolink"
  ),
  badge_cran_download(type = "grand-total"),
  badge_github_actions(action = "R-CMD-check.yaml"),
  badge_codecov(),
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  ),
  paste0(
    "[![",
    "FAIR checklist badge",
    "](",
    "https://img.shields.io/badge/",
    "fairsoftwarechecklist.net", "-",
    "-",
    "00a7d9",
    ")](",
    "https://fairsoftwarechecklist.net/v0.2/?f=31&a=32113&i=32322&r=133",
    ")"
  ),
  paste0(
    "[![",
    "fair-software.eu",
    "](",
    "https://img.shields.io/badge/",
    "fair--software.eu", "-",
    "%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8F",
    "-",
    "green",
    ")](",
    "https://fair-software.eu",
    ")"
  ),
  badge_custom(
    x = "Contributor%20Covenant",
    y = "3.0",
    color = "4baaaa",
    url = "https://www.contributor-covenant.org/version/3/0/code_of_conduct/",
    alt = "Contributor Covenant 3.0 Code of Conduct"
  )
)
```
<!-- badges: end -->

## Overview

`logolink` is an [R](https://www.r-project.org/) package that simplifies setting up and running [NetLogo](https://www.netlogo.org/) simulations directly from R. It offers a modern, streamlined interface for running models, following the [tidyverse principles](https://tidyverse.tidyverse.org/articles/manifesto.html) and integrating seamlessly with the broader [tidyverse ecosystem](https://www.tidyverse.org/).

The package is designed to work with NetLogo **7.0.1 and above**. If you need support for earlier NetLogo releases (up to 7.0.0), use an older version of this package.

> If you find this project useful, please consider giving it a star! &nbsp; [![GitHub Repository Stars](https://img.shields.io/github/stars/danielvartan/logolink)](https://github.com/danielvartan/logolink/)

> The continuous development of `logolink` depends on community support. If you find this project useful, and can afford to do so, please consider becoming a sponsor. &nbsp; [![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

## Another R Package for NetLogo?

While other R packages connect R and NetLogo, `logolink` is currently the only one that fully supports the latest NetLogo release. It is actively maintained, follows tidyverse conventions, and is designed to be simple and straightforward to use.

For context, [`RNetLogo`](https://CRAN.R-project.org/package=RNetLogo) works only with older versions (up to version 6.0.0, released in December 2016) and has not been updated since June 2017. [`nlrx`](https://CRAN.R-project.org/package=nlrx) provides a powerful framework for managing experiments and results, but [supports](https://docs.ropensci.org/nlrx/reference/supported_netlogo_versions.html) only up to NetLogo 6.3.0 (released in September 2022) and has [many unresolved issues](https://github.com/ropensci/nlrx/issues). `logolink` complements these packages by focusing on simplicity, full compatibility with NetLogo 7, and seamless integration into modern R workflows.

## Installation

You can install the released version of `logolink` from [CRAN](https://CRAN.R-project.org) with:

```r
install.packages("logolink")
```

And the development version from [GitHub](https://github.com/) with:

```r
# install.packages("remotes")
remotes::install_github("danielvartan/logolink")
```

## Usage

`logolink` usage is very straightforward. The main functions are:

- [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html): Create a NetLogo BehaviorSpace experiment XML file.
- [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html): Run a NetLogo BehaviorSpace experiment.

Along with this package, you will also need NetLogo 7.0.1 or higher installed on your computer. You can download it from the [NetLogo website](https://www.netlogo.org).

### Setting the NetLogo Path

`logolink` will try to find out the path to the NetLogo installation automatically, but **if it fails**, you will need to set it manually.

This requires the path to the NetLogo installation to be set as an environment variable named `NETLOGO_HOME` when running simulations. The exact path varies depending on your operating system but is usually easy to find. On Windows, for example, it typically looks like `C:\Program Files\NetLogo 7.0.2`.

You can set this environment variable temporarily in your R session using `Sys.setenv("NETLOGO_HOME" = "[PATH]")`, or permanently by adding it to your [`.Renviron`](https://rstats.wtf/r-startup.html#renviron) file.

Example (Windows):

```{r}
#| eval: false

Sys.setenv(NETLOGO_HOME = file.path("C:", "Program Files", "NetLogo 7.0.2"))

Sys.getenv("NETLOGO_HOME")
#> [1] "C:\Program Files\NetLogo 7.0.2"
```

### Creating an Experiment

To start running your model from R you first need to setup an experiment. You can do this by setting a [BehaviorSpace](https://docs.netlogo.org/behaviorspace.html) experiment with the [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html) function. This function will create a [XML](https://en.wikipedia.org/wiki/XML) file that contains all the information about your experiment, including the parameters to vary, the metrics to collect, and the number of runs to perform.

Alternatively, you can set up your experiment directly in NetLogo and save it as part of your model. In this case, you can skip the [`create_experiment`](https://danielvartan.github.io/logolink/reference/create_experiment.html) step and just provide the name of the experiment when running the model with [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html).

Example:

```{r}
library(logolink)
```

```{r}
#| label: setup_file

setup_file <- create_experiment(
  name = "Wolf Sheep Simple Model Analysis",
  repetitions = 10,
  sequential_run_order = TRUE,
  run_metrics_every_step = TRUE,
  setup = "setup",
  go = "go",
  time_limit = 1000,
  metrics = c(
    'count wolves',
    'count sheep'
  ),
  run_metrics_condition = NULL,
  constants = list(
    "number-of-sheep" = 500,
    "number-of-wolves" = list(
      first = 5,
      step = 1,
      last = 15
    ),
    "movement-cost" = 0.5,
    "grass-regrowth-rate" = 0.3,
    "energy-gain-from-grass" = 2,
    "energy-gain-from-sheep" = 5
  )
)
```

```{r}
setup_file |> inspect_experiment_file()
```

### Running an Experiment

With the experiment file created, you can now run your model using the [`run_experiment`](https://danielvartan.github.io/logolink/reference/run_experiment.html) function. This function will execute the NetLogo model with the specified parameters and return the results as a tidy data frame.

```{r}
model_path <-
  Sys.getenv("NETLOGO_HOME") |>
  file.path(
    "models",
    "IABM Textbook",
    "chapter 4",
    "Wolf Sheep Simple 5.nlogox"
  )
```

```{r}
#| label: run_experiment
#| output: false

results <-
  model_path |>
  run_experiment(setup_file = setup_file)
```

```{r}
library(dplyr)

results |> glimpse()
```

### Analyzing the Data (Bonus Section)

Below is a simple example of how to visualize the results using [`ggplot2`](https://ggplot2.tidyverse.org/).

```{r}
#| output: false

library(dplyr)

data <-
  results |>
  summarize(
    across(everything(), ~ mean(.x, na.rm = TRUE)),
    .by = c(step, number_of_wolves)
  ) |>
  arrange(number_of_wolves, step)
```

```{r}
#| label: readme-wolf-sheep-model-plot

library(ggplot2)

data |>
  mutate(number_of_wolves = as.factor(number_of_wolves)) |>
  ggplot(
    aes(
      x = step,
      y = count_sheep,
      group = number_of_wolves,
      color = number_of_wolves
    )
  ) +
    labs(
      x = "Time step",
      y = "Average number of sheep",
      color = "Wolves"
    ) +
  geom_line()
```

Please refer to the [BehaviorSpace Guide](https://docs.netlogo.org/behaviorspace.html) for complete guidance on how to set and run experiments in NetLogo. To gain a better understand of the mechanics behind R and NetLogo communication, see the [Running from the Command Line](https://docs.netlogo.org/behaviorspace.html#running-from-the-command-line) section.

Click [here](https://danielvartan.github.io/logolink/reference/) to see `logolink` full list of functions.

## Citation

If you use this package in your research, please cite it to acknowledge the effort put into its development and maintenance. Your citation helps support its continued improvement.

```{r}
citation("logolink")
```

## License

```{r}
#| echo: false
#| output: asis

cat(
  badge_license(
    license = "GPLv3",
    color = "bd0000",
    url = "https://www.gnu.org/licenses/gpl-3.0"
  )
)
```

```text
Copyright (C) 2025 Daniel Vartanian

logolink is free software: you can redistribute it and/or modify it under the
terms of the GNU General Public License as published by the Free Software
Foundation, either version 3 of the License, or (at your option) any later
version.

This program is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the GNU General Public License for more details.

You should have received a copy of the GNU General Public License along with
this program. If not, see <https://www.gnu.org/licenses/>.
```

## Contributing

```{r}
#| echo: false
#| output: asis

cat(
  badge_custom(
    x = "Contributor%20Covenant",
    y = "3.0",
    color = "4baaaa",
    url = "https://www.contributor-covenant.org/version/3/0/code_of_conduct/",
    alt = "Contributor Covenant 3.0 Code of Conduct"
  )
)
```

Contributions are always welcome! Whether you want to report bugs, suggest new features, or help improve the code or documentation, your input makes a difference. Before opening a new issue, please take a moment to review our [Guidelines for Contributing](https://danielvartan.github.io/logolink/CONTRIBUTING.html) and check the [issues tab](https://github.com/danielvartan/logolink/issues) to see if your topic has already been reported.

[![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

You can also support the development of `logolink` by becoming a sponsor. Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please mention `logolink` in your donation message.

## Acknowledgments

`logolink` brand identity is based on the [NetLogo](https://www.netlogo.org/) brand identity.
